Parallel Hierarchical Clustering using \\ Rank-Two Nonnegative Matrix Factorization \thanks{Identify applicable funding agency here. If none, delete this.}

### 2020-07-16 latex/tex/experiments.tex
hierarchical scaling

From equation \ref{eq:treecost}, we expect to see perfect strong scaling in a computationally bound H2NMF problem with fixed frontier node $k$ count.
However, the problem should become latency bound with its $\log p$ cost as we scale. Figure \ref{fig:fig:synhierspeedup} shows this latency limit as
we scale to 50 compute nodes on the synthetic data. Figure \ref{fig:synhierstrongscaling}

From equation \ref{eq:levelcost}
\begin{itemize}
	\item theory (\Cref{eq:treecost}) says...
	\item Fig 7 shows limit of strong scaling for synthetic data (time increases at 50 nodes compared to 40), Fig 8 gives breakdown and explains
	\begin{itemize}
		\item drastic performance dropoff may be artifact of all-gather, which should take about the same time as reduce-scatter, but speedup from 30 to 40 nodes is also slight
		\item by 50 nodes, about 50\% of time in communication; still mostly dominated by bandwidth cost because ag and rs dominate ar
		\item note the difference between the scaling of single rank-2 and entire tree: poor scaling must be due to lower levels, which we explore in the next section
	\end{itemize}
	\item Fig 9 shows reasonable scaling for entire tree -- speedup drops from 7x to 6x over 80 nodes; Fig 10 shows that MM is still 70\% of run time on 80 nodes, which allows for near-linear scaling
\end{itemize}

>>>>

From equation \ref{eq:treecost}, we expect to see perfect strong scaling in a computationally bound hierarchical clustering problem with fixed frontier node $k=100$ count.
The problem should become latency bound with its $\log p$ cost as we scale.

Figure \ref{fig:synhierstrongscaling} shows the limits of scaling as the hierarchical cost is becoming saturated with communication.
The AllReduce segment would consistently grow if the problem was becoming latency bound, which is not the case.
So, the problem is becoming bandwidth bound as seen through the ReduceScatter and AllGather segments. 
However, computation still makes up more than 60\% of the total cost.

With the much larger \image{} dataset, it's possible to scale much further as seen in Figure \ref{fig:rwhierspeedup} without being overwhelmed by the communication cost.
From Figure \ref{fig:fig:rwhierstrongscaling}, the communication cost doesn't make up more than 20\% of the total cost even scaling to 80 compute nodes.

Note that in order to compare hierarchical clusters between differing compute nodes, we limited the output to include only complete levels in the tree.

\begin{itemize}
	\item theory (\Cref{eq:treecost}) says...
	\item Fig 7 shows limit of strong scaling for synthetic data (time increases at 50 nodes compared to 40), Fig 8 gives breakdown and explains
	\begin{itemize}
		\item drastic performance dropoff may be artifact of all-gather, which should take about the same time as reduce-scatter, but speedup from 30 to 40 nodes is also slight
		\item by 50 nodes, about 50\% of time in communication; still mostly dominated by bandwidth cost because ag and rs dominate ar
		\item note the difference between the scaling of single rank-2 and entire tree: poor scaling must be due to lower levels, which we explore in the next section
	\end{itemize}
	\item Fig 9 shows reasonable scaling for entire tree -- speedup drops from 7x to 6x over 80 nodes; Fig 10 shows that MM is still 70\% of run time on 80 nodes, which allows for near-linear scaling
\end{itemize}
